---
title: "Census Demographic Cleaning"
output: html_notebook
Project: Squeeze Project
Created By: Mirit Friedman
Version Date: 5/18/2022
---

```{r}
#Load packages: tidyverse, dplyr
library(tidyverse)
library(dplyr) ## load the library tidyverse
library(readtext)
```

```{r}
setwd("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads")
```

Read in all the individual files.
```{r}
csd2000 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2000csd.csv")
csd2009 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2009csd.csv")
csd2010 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2010csd.csv")
csd2011 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2011csd.csv")
csd2012 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2012csd.csv")
csd2013 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2013csd.csv")
csd2014 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2014csd.csv")
csd2015 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2015csd.csv")
csd2016 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2016csd.csv")
csd2017 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2017csd.csv")
csd2018 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2018csd.csv")

p2000 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2000p.csv")
p2009 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2009p.csv")
p2010 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2010p.csv")
p2011 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2011p.csv")
p2012 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2012p.csv")
p2013 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2013p.csv")
p2014 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2014p.csv")
p2015 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2015p.csv")
p2016 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2016p.csv")
p2017 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2017p.csv")
p2018 <- read.csv("/Users/miritfriedman/Desktop/Squeeze Data/census_demographics/CensusDownloads/cen_2018p.csv")
```

Subset the data columns of interest.
```{r}
csd2000<-csd2000%>%select(c(9,14,23,24,55,71,74,78,79,81,86,91))
p2000<-p2000%>%select(c(9,14,23,24,55,71,74,78,79,81,86,91))
csd2009<-csd2009%>%select(c(1,54,63,64,95,111,114,118,119,121,126,131))
p2009<-p2009%>%select(c(1,54,63,64,95,111,114,118,119,121,126,131))
csd2010<-csd2010%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2010<-p2010%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2011<-csd2011%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2011<-p2011%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2012<-csd2012%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2012<-p2012%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2013<-csd2013%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2013<-p2013%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2014<-csd2014%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2014<-p2014%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2015<-csd2015%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2015<-p2015%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2016<-csd2016%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2016<-p2016%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2017<-csd2017%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2017<-p2017%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
csd2018<-csd2018%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
p2018<-p2018%>%select(c(1,56,65,66,97,113,116,120,121,123,128,133))
```

Change the column names and remove the first row.
```{r}
colnames(csd2000)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2000<-csd2000[-c(1),]

colnames(csd2009)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2009<-csd2009[-c(1),]

colnames(csd2010)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2010<-csd2010[-c(1),]

colnames(csd2011)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2011<-csd2011[-c(1),]

colnames(csd2012)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2012<-csd2012[-c(1),]

colnames(csd2013)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2013<-csd2013[-c(1),]

colnames(csd2014)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2014<-csd2014[-c(1),]

colnames(csd2015)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2015<-csd2015[-c(1),]

colnames(csd2016)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2016<-csd2016[-c(1),]

colnames(csd2017)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2017<-csd2017[-c(1),]

colnames(csd2018)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
csd2018<-csd2018[-c(1),]

colnames(p2000)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2000<-p2000[-c(1),]

colnames(p2009)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2009<-p2009[-c(1),]

colnames(p2010)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2010<-p2010[-c(1),]

colnames(p2011)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2011<-p2011[-c(1),]

colnames(p2012)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2012<-p2012[-c(1),]

colnames(p2013)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2013<-p2013[-c(1),]

colnames(p2014)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2014<-p2014[-c(1),]

colnames(p2015)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2015<-p2015[-c(1),]

colnames(p2016)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2016<-p2016[-c(1),]

colnames(p2017)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2017<-p2017[-c(1),]

colnames(p2018)<-c("FIPS","TOT_POP","PCT_White","PCT_Black","PCT_Hispanic","PCT_25Plus_LessThanDiploma","PCT_25Plus_Bachelors","Median_Income","Median_Year_Built","Below18_Poverty","18_64_Poverty","65Plus_Poverty")
p2018<-p2018[-c(1),]

```

Change the variables from character to numeric.
```{r}
csd2000<-csd2000 %>% mutate_if(is.character,as.numeric)
csd2009<-csd2009 %>% mutate_if(is.character,as.numeric)
csd2010<-csd2010 %>% mutate_if(is.character,as.numeric)
csd2011<-csd2011 %>% mutate_if(is.character,as.numeric)
csd2012<-csd2012 %>% mutate_if(is.character,as.numeric)
csd2013<-csd2013 %>% mutate_if(is.character,as.numeric)
csd2014<-csd2014 %>% mutate_if(is.character,as.numeric)
csd2015<-csd2015 %>% mutate_if(is.character,as.numeric)
csd2016<-csd2016 %>% mutate_if(is.character,as.numeric)
csd2017<-csd2017 %>% mutate_if(is.character,as.numeric)
csd2018<-csd2018 %>% mutate_if(is.character,as.numeric)

p2000<-p2000 %>% mutate_if(is.character,as.numeric)
p2009<-p2009 %>% mutate_if(is.character,as.numeric)
p2010<-p2010 %>% mutate_if(is.character,as.numeric)
p2011<-p2011 %>% mutate_if(is.character,as.numeric)
p2012<-p2012 %>% mutate_if(is.character,as.numeric)
p2013<-p2013 %>% mutate_if(is.character,as.numeric)
p2014<-p2014 %>% mutate_if(is.character,as.numeric)
p2015<-p2015 %>% mutate_if(is.character,as.numeric)
p2016<-p2016 %>% mutate_if(is.character,as.numeric)
p2017<-p2017 %>% mutate_if(is.character,as.numeric)
p2018<-p2018 %>% mutate_if(is.character,as.numeric)
```

Create a column with the total percent below the poverty line and delete counts of people below poverty line. 
```{r}
csd2000<-csd2000 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2009<-csd2009 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2010<-csd2010 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2011<-csd2011 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2012<-csd2012 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2013<-csd2013 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2014<-csd2014 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2015<-csd2015 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2016<-csd2016 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2017<-csd2017 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
csd2018<-csd2018 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))

p2000<-p2000 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2009<-p2009 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2010<-p2010 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2011<-p2011 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2012<-p2012 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2013<-p2013 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2014<-p2014 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2015<-p2015 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2016<-p2016 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2017<-p2017 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
p2018<-p2018 %>% rowwise() %>% mutate(POV_POP = sum(c_across(10:12)))
```

Add a column with the data year.
```{r}
csd2000$Year<-2000
csd2009$Year<-2009
csd2010$Year<-2010
csd2011$Year<-2011
csd2012$Year<-2012
csd2013$Year<-2013
csd2014$Year<-2014
csd2015$Year<-2015
csd2016$Year<-2016
csd2017$Year<-2017
csd2018$Year<-2018

p2000$Year<-2000
p2009$Year<-2009
p2010$Year<-2010
p2011$Year<-2011
p2012$Year<-2012
p2013$Year<-2013
p2014$Year<-2014
p2015$Year<-2015
p2016$Year<-2016
p2017$Year<-2017
p2018$Year<-2018
```

Merge the individual datasets into one file.
```{r}
merge<- rbind(csd2000,csd2009,csd2010,csd2011,csd2012,csd2013,csd2014,csd2015,csd2016,csd2017,csd2018,p2000,p2009,p2010,p2011,p2012,p2013,p2014,p2015,p2016,p2017,p2018)
```

Read in the master list.
```{r}
masterFIPS<-read.csv("/Users/miritfriedman/Desktop/Squeeze Data/FinalList_FIPSKey.csv")
masterFIPS<-rename(masterFIPS, "FIPS"="FIPS_2020")
```

Subset the census data by the master list.
```{r}
merge<-as.data.frame(merge)
census_output<- filter(merge, FIPS %in% masterFIPS$FIPS)
```

Create a column with the percent below the poverty line
```{r}
census_output$POV_PCT<-(census_output$POV_POP/census_output$TOT_POP)*100
```

Count unique outputs to verify how many FIPS from the master list had census data.
```{r}
uni<-unique(census_output$FIPS)
uniFIP<-unique(masterFIPS$FIPS)
uniPWS<-unique(masterFIPS$PWSID)
dup<-masterFIPS[duplicated(masterFIPS$FIPS),]
dup2<-masterFIPS[duplicated(masterFIPS$PWSID),]
```

Save the output file
```{r}
date<-Sys.Date()
csvfilename<-paste("CensusDem_Output_",date,".csv",sep="")
write.csv(census_output,file=csvfilename)
```

Descriptive Statistics
```{r}
summary(census_output$PCT_White)
summary(census_output$Median_Year_Built)
summary(census_output$POV_PCT)
summary(census_output$TOT_POP)
summary(census_output$Median_Income)
```




Create list of FIPS to match county voting data

```{r}
cnty<-subset(csd2000, select= c(2,9,10,12))
```

